Automated Metadata Generation for Fish Specimen Image Collections

### 2021-03-27 paper/paper.tex
Added all raw results (no tables or pictures yet)
Also updated validate.py to show a few more things

%% A "teaser" image appears between the author and affiliation
%% information and the body of the document, and typically spans the
%% page.
\begin{teaserfigure}
  \centering
  \begin{subfigure}%{.5\linewidth}
  \includegraphics[width=.49\textwidth]{images/teaser1}
  \end{subfigure}
  \begin{subfigure}%{.5\linewidth}
  \includegraphics[width=.49\textwidth]{images/teaser2}
  \end{subfigure}
  \caption{Initial object detection on specimen images using Detectron2~\cite{wu2019detectron2}}
  \Description{Fish predictions}
  \label{fig:fish}
\end{teaserfigure}

@@

\section{Goal and Objectives}

@@

For the fish \verb|length| property, it is necessary to determine the furthest points from the centroid of the fish in each direction. Since fish are normally measured in a straight line from their snout down the middle of their trunk, we project every pixel onto the major axis of the fish (as a line through its centroid) before checking the distance. This projection is done by intersecting the line through the point along the tangent of the principal axis and the aforementioned centroid--principal axis line. After checking every pixel, measuring the distance between the two furthest projected points gives the length of the fish in pixels. Multiplying this by \verb|scale| gives the fish \verb|length| in centimeters.
\begin{comment}
\subsection{Consideration}
Just gonna throw some caveats in here, might be its own section might not
\begin{itemize}
    \item Program falls back to using detectron mask and bbox for metadata generation in the event the mask generation/pixel analysis fails. Failure is defined as never finding a cohesive flood filled mask that covers more than 10\% of the detectron bbox initially. Can provide a statistic on how often this occurs.
\end{itemize}
\end{comment}
\section{Results}
Technicians employed by our team have manually generated the 22 metadata properties deemed crucial to the overall BGNN project~\cite{Leipzig2021.01.28.428644} for a large number of INHS images. \(20,699\) total entries were created by 13 technicians that spanned \(8,398\) unique images. Of these \(8,398\) images, \(7,247\) were both not part of the \verb|detectron| training set and meet our current admissibility criteria for what \verb|detectron| and the pixel analysis process can reliably handle. We ran the metadata extraction program on these \(7,247\) images.

Our automated process currently generates 7 of the 22 core metadata properties: \verb|if_fish| (\verb|has_fish|), \verb|fish_number| (\verb|fish_count|), \verb|if_ruler| (\verb|has_ruler|), \verb|specimen_angled| (\verb|clock_value|), \verb|specimen_view| (\verb|side|), \verb|brightness| (derived from \verb|background.mean| and \verb|foreground.mean|), and \verb|if_background_uniform| (\verb|background.std|). In addition, our process also calculates bounding boxes and fish lengths in centimeters.\todo{Talk more about what we can and can't verify and how. Also, need to fit the percent of the time pixel analysis failed in here somewhere.}

All images in the INHS dataset contain exactly one fish. \textbf{Todo}: if\_fish and fish\_count results

\textbf{Todo}: Ruler results

\textbf{Todo}: Side results

\textbf{Todo}: Clock value results

\textbf{Todo}: Brightness/uniformity results? I can try to do this but I need to set up the infrastructure, we probably need to discuss a plan of action for this.

\textbf{Todo}: Bounding box results

\textbf{Todo}: Scale results
\begin{comment}
\begin{verbatim}
What I think should be the final results:

>>>>

%% A "teaser" image appears between the author and affiliation
%% information and the body of the document, and typically spans the
%% page.
\begin{teaserfigure}
  \centering
  \includegraphics[width=.49\textwidth]{images/teaser1}
  \includegraphics[width=.49\textwidth]{images/teaser2}
  \caption{Initial object detection on specimen images using Detectron2~\cite{wu2019detectron2}}
  \Description{Fish predictions}
  \label{fig:fish}
\end{teaserfigure}

@@

\section{Goals and Objectives}

@@

For the fish \verb|length| property, it is necessary to determine the furthest points from the centroid of the fish in each direction. Since fish are normally measured in a straight line from their snout down the middle of their trunk, we project every pixel onto the major axis of the fish (as a line through its centroid) before checking the distance. This projection is done by intersecting the line through the point along the tangent of the principal axis and the aforementioned centroid--principal axis line. After checking every pixel, measuring the distance between the two furthest projected points gives the length of the fish in pixels. Multiplying this by \verb|scale| gives the fish \verb|length| in centimeters.
\begin{comment}
\subsection{Consideration}
Just gonna throw some caveats in here, might be its own section might not
\begin{itemize}
    \item Program falls back to using detectron mask and bbox for metadata generation in the event the mask generation/pixel analysis fails. Failure is defined as never finding a cohesive flood filled mask that covers more than 10\% of the detectron bbox initially. Can provide a statistic on how often this occurs.
\end{itemize}
\end{comment}
\section{Results}
Technicians employed by our team have manually generated the 22 metadata properties deemed crucial to the overall BGNN project~\cite{Leipzig2021.01.28.428644} for a large number of INHS images. \(20,699\) total entries were created by 13 technicians that spanned \(8,398\) unique images. Of these \(8,398\) images, \(7,244\) were both not part of the \verb|detectron| training set and meet our current admissibility criteria for what \verb|detectron| and the pixel analysis process can reliably handle. We ran the metadata extraction program on these \(7,244\) images.

Our automated process currently generates 7 of the 22 core metadata properties: \verb|if_fish| (\verb|has_fish|), \verb|fish_number| (\verb|fish_count|), \verb|if_ruler| (\verb|has_ruler|), \verb|specimen_angled| (\verb|clock_value|), \verb|specimen_view| (\verb|side|), \verb|brightness| (derived from \verb|background.mean| and \verb|foreground.mean|), and \verb|if_background_uniform| (\verb|background.std|). In addition, our process also calculates bounding boxes and fish lengths in centimeters.\todo{Talk more about what we can and can't verify and how.}

\subsection{Fish Detection}
All images in the INHS dataset contain exactly one fish. For \(7,209\) of the specimen images, one fish was detected. For 25 of the images, 2 fish were detected, 3 fish were detected for 3 images, and for 7 images no fish were found. In the case of greater than 1 fish, 9 of the 28 contained tags which overlapped the fish and were themselves labeled as a second fish. Of the remaining 17, \verb|detectron| erroneously labeled the fish as two separate fish objects, or labeled a subsection of the fish a second time.

\subsection{Ruler Detection}
For all but 2 of the \(7,244\) images \verb|detectron| was able to find the ruler.\todo{Not sure if I can phrase it that way or not since I haven't looked at all the images to see if the ruler itself was found or something else was labeled a ruler} For an additional 56 images, the ruler itself was found but the ``2'' and/or the ``3'' on it was not, so a scale calculation could not be performed.

\subsection{Side Detection}
For 246 of the images \verb|detectron| was unable to find a fish eye. Of the remaining \(6,998\) images where it did, the correct side (\verb|left| or \verb|right|) was detected in all but 6 cases.

\subsection{Clock Value}
Clock position values were successfully generated for \(6,991\) of the images. Of those, all but 10 were within \(\pm{}1\) of the correct result.

\subsection{Brightness/Uniformity Results?}
\textbf{Todo}: I can try to do this but I need to set up the infrastructure, we probably need to discuss a plan of action for this.

\subsection{Bounding Box}
Fish bounding boxes were calculated for all \(7,237\) images in which a fish was found. All but 263 of these were generated via pixel analysis, with those 263 falling back to the original \verb|detectron| bounding box. \textbf{Todo}: Anything more than this is going to need hand checked.

\subsection{Length}
Fish lengths were calculated for \(7,179\) of the images. \textbf{Todo}: Anything more than this is going to need hand checked.
\begin{comment}
\begin{verbatim}
What I think should be the final results:
